---
title: "Homework 2"
author: "Charlotte Fowler"
date: "10/01/2019"
output: github_document
---

```{r}
library(tidyverse)
library(readxl)
```


# Problem 1
```{r}
#reading in relevant data from sheet 1
trash_wheel = 
  read_excel("data/Trash-Wheel-Collection-Totals-8-6-19.xlsx", 
                         sheet = 'Mr. Trash Wheel',
                         range = "A2:N338") %>% 
  janitor::clean_names() %>%  
  drop_na(dumpster)

#reading in 2017 data
precip_2017 =
  read_excel("data/Trash-Wheel-Collection-Totals-8-6-19.xlsx", 
                         sheet = '2017 Precipitation',
                         range = "A2:B15") %>%
  janitor::clean_names() %>% 
  drop_na('month', 'total') %>% 
  mutate(year = 2017)

#reading in 2018 data
precip_2018 =
  read_excel("data/Trash-Wheel-Collection-Totals-8-6-19.xlsx", 
                         sheet = '2018 Precipitation',
                         range = "A2:B15") %>%
  janitor::clean_names() %>% 
  drop_na('month','total') %>% 
  mutate(year = 2018)
```

```{r}
#combining 2017 and 2018 data
precipitation = 
  bind_rows(precip_2017, precip_2018) %>% 
  mutate(month = factor(month, labels=c(month.name))) 
```


The Mr. Trash Wheel data has `r nrow(trash_wheel)` observations and `r ncol(trash_wheel)` variables, while the precipiation data from 2017 and 2018 has `r nrow(precipitation)` observations and `r ncol(precipitation)` variables. Mr. Trash Wheel provides information on the `dumpster`, `month`, and `year` of the trash collection, as well as information on the quantity and content of the trash. The precipitation data only includes information on the `month`, `year`, and `precipitation` total. 
In 2018, the total precipitation recorded was `r sum(pull(precip_2018, total))`. 
The median number of sports balls per dumster in 2017 was `r filter(trash_wheel, year == 2017) %>% summarize(median(sports_balls))`. 


# Problem 2


```{r}
#reading and cleaning pols_month
pols_month = 
  read_csv("data/pols-month.csv") %>% 
  separate(mon, into = c("year", "month", "day")) %>% 
  mutate(month = factor(month, labels=c(month.name))) %>% 
  mutate(president = ifelse(prez_gop == 1, "republican", "democrat")) %>% 
  select(-prez_dem, -prez_gop, -day)
```

```{r}
#reading in and cleaning snp data
snp = 
  read_csv("data/snp.csv") %>% 
  separate(date, into = c("month", "day", "year")) %>% 
  arrange(month) %>% 
  mutate(month = factor(month, labels = c(month.name))) %>% 
  select(year, month, close) %>% 
  arrange(year, month)
```


```{r}
#reading in and cleaning unemployment data
unemployment = 
  read_csv("data/unemployment.csv") %>% 
  rename(year = Year, January = Jan, February = Feb, March = Mar, April = Apr, June = Jun, July = Jul, August = Aug, September = Sep, October = Oct, November = Nov, December = Dec) %>% 
  pivot_longer(
   January:December,
   names_to = "month",
   values_to = "unemployment"
  ) %>% 
  mutate(year = factor(year))
```

```{r}
#merging snp and pols_month
data_538 = 
  left_join(pols_month, snp, by= c('year','month'))

#merging data_538 and unemployment
data_538 = 
  left_join(data_538, unemployment, by= c('year','month'))
```


First, the `r nrow(pols_month)` rows by `r ncol(pols_month)` variables `pols-month` dataset contains information by date about the United States government's distribution between democrats and republicans in the House, Senate, and the president. Next, the `r nrow(snp)` rows by `r ncol(snp)` variables `snp` dataset gives information by date on the S&P stock market value. Lastly, the `r nrow(unemployment)` rows by `r ncol(unemployment)` variables `unemployment` dataset gives information by month and year on the percentage of unemployed individuals in the country. 
Together, merged into `data_538`, we have a `r nrow(data_538)` rows by `r ncol(data_538)`columns dataset with information from `r min(pull(data_538, year))` to `r max(pull(data_538, year))`. This dataset has information organized by `year` and `month` about the distribution of democrats and republicans in the Senate and the House, the party of the president (`president`), the close value of the S&P stock price (`close`), and the unemployment rate (`unemployment`).






# Problem 3

```{r}
#reading in and cleaning baby names data
popular_baby_names = 
  read_csv("data/Popular_Baby_Names.csv") %>% 
  janitor::clean_names() %>% 
  mutate(
    ethnicity = replace(ethnicity, ethnicity == "ASIAN AND PACIFIC ISLANDER", "ASIAN AND PACI"),
    ethnicity = replace(ethnicity, ethnicity == "BLACK NON HISPANIC", "BLACK NON HISP"),
    ethnicity = replace(ethnicity, ethnicity == "WHITE NON HISPANIC", "WHITE NON HISP"),
    childs_first_name = str_to_lower(childs_first_name),
    ethnicity = str_to_lower(ethnicity)
    ) %>% 
  distinct()
```
```{r}
#creating new dataframe for Olivia only, printing table 
olivia_table = popular_baby_names %>% 
  filter(childs_first_name == 'olivia') %>% 
  select (year_of_birth, ethnicity, rank) %>% 
  pivot_wider(
    names_from = year_of_birth,
    values_from = rank
  ) %>% 
  mutate(
    ethnicity = replace(ethnicity, ethnicity == "asian and paci", "Asian and Pacific Islander"),
    ethnicity = replace(ethnicity, ethnicity == "black non hisp", "Black and Non Hispanic"),
    ethnicity = replace(ethnicity, ethnicity == "white non hisp", "White and Non Hispanic"), 
    ethnicity = replace(ethnicity, ethnicity == "hispanic", "Hispanic")
  ) %>% 
  rename(Ethnicity = ethnicity)

knitr::kable(olivia_table)
```


```{r}
#creating new dataframe for most popular boys names, printing table
boys_table = popular_baby_names %>% 
  filter(
    gender == 'MALE',
    rank == 1
    ) %>% 
  select(year_of_birth, ethnicity, childs_first_name) %>% 
  mutate(childs_first_name = str_to_title(childs_first_name)) %>% 
  pivot_wider(
    names_from = year_of_birth,
    values_from = childs_first_name
  ) %>% 
  mutate(
    ethnicity = replace(ethnicity, ethnicity == "asian and paci", "Asian and Pacific Islander"),
    ethnicity = replace(ethnicity, ethnicity == "black non hisp", "Black and Non Hispanic"),
    ethnicity = replace(ethnicity, ethnicity == "white non hisp", "White and Non Hispanic"),
    ethnicity = replace(ethnicity, ethnicity == "hispanic", "Hispanic")
  ) %>% 
  rename(Ethnicity = ethnicity)  
  

knitr::kable(boys_table)
```


```{r}
#Creating scatterplot for white non-hisp boys names in 2016
popular_baby_names %>% 
  filter(
    gender == 'MALE',
    ethnicity == 'white non hisp',
    year_of_birth == 2016
    ) %>% 
  ggplot(aes(y = count, x = rank)) +
  geom_point() + 
  labs(
    title = 'White Non-Hispanic Boys Names in 2016',
    x = "Name Rank",
    y = "Name Frequency"
  )
```





