---
title: "Homework 2"
author: "Charlotte Fowler"
date: "9/28/2019"
output: github_document
---

```{r}
library(tidyverse)
library(readxl)
```


# Problem 1
```{r}
#reading in relevant data from sheet 1
trash_wheel = 
  read_excel("data/HealthyHarborWaterWheelTotals2018-7-28.xlsx", 
                         sheet = 'Mr. Trash Wheel',
                         range = "A2:N338") %>% 
  #creating better variable names
  janitor::clean_names() %>%  
  #drop rows with na in dumpster
  drop_na(dumpster)

#reading in 2017 data
precip_2017 =
  read_excel("data/HealthyHarborWaterWheelTotals2018-7-28.xlsx", 
                         sheet = '2017 Precipitation',
                         range = "A2:B15") %>%
  janitor::clean_names() %>% 
  drop_na('month') %>% 
  mutate(year = 2017)

#reading in 2018 data
precip_2018 =
  read_excel("data/HealthyHarborWaterWheelTotals2018-7-28.xlsx", 
                         sheet = '2018 Precipitation',
                         range = "A2:B15") %>%
  janitor::clean_names() %>% 
  drop_na('month','total') %>% 
  mutate(year = 2018)
```

```{r}
#combining 2017 and 2018 data
precipitation = 
  bind_rows(precip_2017, precip_2018) %>% 
  mutate(month = factor(month, labels=c(month.name))) 
```


The Mr. Trash Wheel data has `r nrow(trash_wheel)` observations, while the precipiation data from 2017 and 2018 has `r nrow(precipitation)` observations. Mr. Trash Wheel provides information on the `dumpster`, `month`, and `year` of the trash collection, as well as information on the quantity and content of the trash. The precipitation data only includes information on the `month`, `year`, and `precipitation` total. 
In 2018, the total precipitation recorded was `r sum(pull(precip_2018, total))`. 
The median number of sports balls per dumster in 2017 was `r filter(trash_wheel, year == 2017) %>% summarize(median(sports_balls))`. 




# Problem 2

First, clean the data in pols-month.csv. Use separate() to break up the variable mon into integer variables year, month, and day; 

replace month number with month name; 

create a president variable taking values gop and dem, 

remove  prez_dem and prez_gop; and remove the day variable
```{r}
#reading and cleaning pols_month
pols_month = 
  read_csv("data/pols-month.csv") %>% 
  separate(mon, into = c("year", "month", "day")) %>% 
  mutate(month = factor(month, labels=c(month.name))) %>% 
  mutate(president = ifelse(prez_gop == 1, "republican", "democrat")) %>% 
  select(-prez_dem, -prez_gop, -day)
```
Second, clean the data in snp.csv using a similar process to the above. For consistency across datasets, arrange according to year and month, and organize so that year and month are the leading columns.

```{r}
#reading in and cleaning snp data
snp = 
  read_csv("data/snp.csv") %>% 
  separate(date, into = c("month", "day", "year")) %>% 
  arrange(month) %>% 
  mutate(month = factor(month, labels=c(month.name))) %>% 
  select(year, month, close) %>% 
  arrange(year, month)
```

tidy the unemployment data so that it can be merged with the previous datasets:
switching from “wide” to “long” format; 
ensuring that key variables have the same name;
ensuring that key variables take the same values.
```{r}
#reading in and cleaning unemployment data
unemployment = 
  read_csv("data/unemployment.csv") %>% 
  rename(year = Year, January = Jan, February = Feb, March = Mar, April = Apr, June = Jun, July = Jul, August = Aug, September = Sep, October = Oct, November = Nov, December = Dec) %>% 
  pivot_longer(
   January:December,
   names_to = "month",
   values_to = "unemployment"
  ) %>% 
  mutate(year = factor(year))
```
Join the datasets by merging snp into pols, and merging unemployment into the result.
```{r}
#merging snp and pols_month
data_538 = 
  left_join(pols_month, snp, by= c('year','month'))

#merging data_538 and unemployment
data_538 = 
  left_join(data_538, unemployment, by= c('year','month'))
```


First, the `pols-month` dataset contains information by date about the United States government's distribution between democrats and republicans in the House, Senate, and the president. Next, the `snp` dataset gives information by date on the S&P stock market value. Lastly, the `unemployment` dataset gives information by month and year on the percentage of unemployed individuals in the country. 
Together, merged into `data_538`, we have a `r nrow(data_538)` rows by `r ncol(data_538)`columns dataset with information from `r min(pull(data_538, year))` to `r max(pull(data_538, year))`. This dataset has information organized by `year` and `month` about the distribution of democrats and republicans in the Senate and the House, the party of the president (`president`), the close value of the S&P stock price (`close`), and the unemployment rate (`unemployment`).

# Problem 3

Load and tidy the data. Note that, although these data may seem fairly well formatted initially, 
the names of a categorical predictor and the case structure of string variables- names - changed over time; you’ll need to address this in your data cleaning. 
Also, some rows seem duplicated, and these will need to be removed (hint: google something like “dplyr remove duplicate rows” to get started).
```{r}
popular_baby_names = 
  read_csv("data/Popular_Baby_Names.csv")
```

```{r}
table(popular_baby_names$Ethnicity)
```

